### 2021-10-23 text/outline/outline.tex
add more text to thresholding theory section

\begin{document}
\noindent
Tim, Gene, Kathryn
\begin{center}
\textbf{CRISPR genome engineering, single-cell sequencing, and exponential family measurement error models}
\end{center}

\begin{abstract}
CRISPR genome engineering and single-cell sequencing have transformed biological discovery. Single-cell CRISPR screens unite these two technologies, linking genetic perturbations in individual cells to changes in gene expression and illuminating regulatory networks underlying diseases. In this work we study single-cell CRISPR screens from a statistical perspective. First, we demonstrate on real data that a standard method for estimation and inference in single-cell CRISPR screens — “thresholded regression” — exhibits attenuation bias and a bias-variance tradeoff as a function of an intrinsic tuning parameter. We recover these phenomena in precise theoretical terms in an idealized Gaussian setting. Next, we introduce GLM-EIV (“generalized linear model with errors-in-variables”), a new method for single-cell CRISPR screen analysis. GLM-EIV generalizes the classical errors-in-variables model to response distributions and sources of measurement error that are exponential family-distributed, overcoming limitations of thresholded regression. We develop a computational infrastructure to deploy GLM-EIV across hundreds or thousands of processors on clouds (e.g., Microsoft Azure) and high-performance clusters. Leveraging this infrastructure, we apply GLM-EIV to analyze two recent, large-scale, single-cell CRISPR screen datasets, yielding new biological insights.
\end{abstract}

@@

\begin{equation}\label{theoretical_model}
\begin{cases}
m_i = \beta^m_0 + \beta^m_1 p_i + \ep_i \\
g_i = \beta^g_0 + \beta^g_1 p_i + \tau_i \\
p_i \sim \textrm{Bern}(\pi) \\
\ep_i, \tau_i \sim N(0,1) \\
p_i \indep \tau_i \indep \ep_i.
\end{cases}
\end{equation}
For a given threshold $c \in \R$, let $\hat{p}_i = \mathbb{I}(g_i \geq c).$ The thresholding estimator $\hat{\beta}^m_1$ for $\beta^m_1$ is given by $$\hat{\beta}^m_1 = \frac{\sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}}) (m_i - \overline{m})}{\sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}})}.$$

\begin{proposition}
 The limit in probability (as $n \to \infty$) of $\hat{\beta}^m_0$ is
\begin{equation}\label{thresh_est_intercepts}
\hat{\beta}^m_1 \xrightarrow{P} \beta^m_1 \left(\frac{ \pi( \omega - \E[ \hat{p}_i ])}{ \E[\hat{p}_i] (1 - \E[\hat{p}_i])}\right),
\end{equation}
where
$$\begin{cases}
\omega = 1 - \Phi\left(c - \beta_1^g - \beta_0^g \right) ,\\ \zeta = 1 - \Phi\left( c - \beta^g_0 \right), \\ \E[\hat{p}_i] = \zeta(1-\pi) + \omega\pi. 
\end{cases}$$
\end{proposition}
Let $\gamma: \R^4 \to \R$ be defined by
$$ \gamma( \beta^g_0, \beta^g_1, c, \pi ) = \frac{\pi (\omega - \E[\hat{p}_i])}{\E[\hat{p}_i] (1 - \E[\hat{p}_i]) }.$$ We call $\gamma$ the ``attenuation function.'' Observe that \begin{itemize}
\item[i.] $\gamma$ does not depend on $\beta^m_1$ or $\beta^m_0$, and
\item[ii.] $\hat{\beta}^m_1 \xrightarrow{P} [\gamma(\beta_0^g, \beta_1^g, c, \pi)] \beta^m_1.$
\end{itemize}
Let $b: \R^4 \to \R$ be the asymptotic bias of $\hat{\beta}^m_1$. We can express $b$ in terms of $\gamma$ as $$b\left(\beta^g_0, \beta^g_1, c, \pi \right) = \beta^m_1 - \gamma( \beta_0^g, \beta_1^g, c, \pi) \beta^m_1 = \beta^m_1 \left[ 1- \gamma(\beta^g_0, \beta^g_1, c, \pi) \right].$$ The asymptotic bias vanishes when the attenuation function equals $1$.

@@

To investigate the basic question of ``What is a good threshold selection strategy?'', we study the relationship between the asymptotic bias $b$ of $\hat{\beta}^m_1$ and the selected threshold $c$. For simplicity, we begin by setting the perturbation probability $\pi$ to $1/2$. Additionally, without loss of generality, we set the target of inference $\beta^m_1$ to $1$. Let $c_\textrm{bayes} \in \R$ be the Bayes-optimal decision boundary for classifying cells as perturbed or unperturbed, i.e. $$c_\textrm{bayes} = \argmin_{c \in \R} \P(\hat{p}_i \neq p_i).$$ Simple algebra shows that $c_\textrm{bayes} = \beta_0^g + (1/2) \beta^g_1.$ Below, we state several results relating the limiting bias $b$ of $\hat{\beta}^m_1$ to $c_\textrm{bayes}$ and $\beta^g_1$. We defer proofs to the appendix.

\begin{itemize}
\item \begin{proposition}
The Bayes-optimal threshold $c_\textrm{bayes}$ is a critical value of the limiting bias $b$ (holding other parameters fixed; Figure \ref{thresholding_theoretical}a, all panels).
\end{proposition}
For certain values of $\beta^g_0$ and $\beta^g_1$, $c_\textrm{bayes}$ is a maximizer of the bias (Figure \ref{thresholding_theoretical}a, left); for other values of $\beta^g_0$ and $\beta^g_1$, $c_\textrm{bayes}$ is a minimizer of the bias (Figure \ref{thresholding_theoretical}a, right).
\item
\begin{proposition}
The limit of the bias in $c$ is $1/2$, i.e. 
$$\lim_{c \to \infty} b(\beta^g_0, \beta^g_1, c, 1/2) = 1/2.$$
\end{proposition}
 In other words, we always can set the threshold to a large number and achieve a bias of $1/2$ (Figure \ref{thresholding_theoretical}a, all panels). This fact establishes an upper bound on the bias of thresholded regression (under optimal threshold selection).
 \item
 \begin{proposition}
 For $\beta^g_1 < 2 \Phi(3/4) \approx 1.35$, $c = c_\textrm{bayes}$ yields a \textit{larger} bias than $c = \infty$ (Figure \ref{thresholding_theoretical}a, left); for $\beta^g_1 > 2 \Phi(3/4)$, $c_\textrm{bayes}$ yields a \textit{smaller} bias than $c = \infty$ (Figure \ref{thresholding_theoretical}a, right); finally, for $\beta^g_1 = 2 \Phi(3/4)$, $c = c_\textrm{bayes}$ and $c = \infty$ both yield a bias of $1/2$ (Figure \ref{thresholding_theoretical}a, middle).
 \end{proposition}
\item \begin{proposition}
Limiting bias decreases monotonically as $\beta_1^g$ increases (Figure \ref{thresholding_theoretical}a; fix threshold and scan panels from left to right).
\end{proposition}
This is consistent with the intuition that the problem becomes easier as the gRNA mixture distribution becomes increasingly well-separated.
\item 
\begin{proposition} For all $\beta^g_0, \beta^g_1,$ and $c \in \R,$ the incurs strict attenuation bias (Figure \ref{thresholding_theoretical}a, all panels).
\end{proposition}
\end{itemize}

These results are subtle, but we can summarize them as follows. First, selecting a threshold that minimizes the bias is deceptively challenging, as there is no rule of thumb that we can apply universally (e.g., ``always choose the Bayes-optimal decision boundary'' or ``always choose a large number''). Second, even if we \textit{have} selected a good threshold, the thresholding method incurs a strict attenuation bias.

\begin{center}
\textbf{Generalizing to $\pi \in [0,1/2]$ (Panel b)}
\end{center}

Next, we generalize the expression for bias to arbitrary $\pi \in [0,1/2]$:
\begin{proposition}
The following relationship holds:
$$ \lim_{ c \to \infty } b(\beta^g_0, \beta^g_1, c, \pi) = \pi.$$
\end{proposition}
In other words, if the perturbation probability is $\pi$, and if we set the threshold to a large number, then the bias also is $\pi$. We can understand this result intuitively by considering an extreme example. When $\pi$ is very small (e.g., $\pi = 0.01$), almost all cells are unperturbed. Therefore, in selecting a large threshold, we correctly classify nearly all unperturbed cells as unperturbed; on the other hand, the \textit{perturbed} cells that we misclassify as \textit{unperturbed} are swamped in number by the truly unperturbed cells, resulting in a small bias.

\begin{figure}[h!]
	\centering
	\includegraphics[width=1\linewidth]{../../figures/thresholding_theoretical/plot}
	\caption{\textbf{Theoretical challenges of thresholded regression.}}
	\label{thresholding_theoretical}
\end{figure}

@@

To shed light on possible limitations of the large threshold selection strategy, we derive an exact bias-variance decomposition for the thresholding estimator. We consider a slightly simpler, no-intercept version of (\ref{theoretical_model}):
\begin{equation}\label{theoretical_model_no_int}
\begin{cases}
m_i = \beta_m p_i + \ep_i \\
g_i = \beta_g p_i + \tau_i \\
p_i \sim \textrm{Bern}(\pi) \\
\ep_i, \tau_i \sim N(0,1) \\
p_i \indep \tau_i \indep \ep_i,
\end{cases}
\end{equation}
The thresholding estimator $\hat{\beta}_m$ is given by
\begin{equation}\label{thresh_estimator_no_int}
 \hat{\beta}_m = \frac{ \sum_{i=1}^n \hat{p}_i m_i }{\sum_{i=1}^n \hat{p}_i^2}. 
\end{equation}
\begin{proposition}
The limiting distribution of $\hat{\beta}_m$ is
$$\sqrt{n}\left(\hat{\beta}_m - l \right) \xrightarrow{d} N\left(0, v\right).$$
\end{proposition}
The bias decreases and the variance increases as the threshold tends to infinity, consistent with the intuition that a large threshold yields a low misclassification rate at the cost of a small ``effective sample size.'' The best strategy for maximizing estimation accuracy (as quantified by mean squared error) is to select a threshold that induces moderate bias. A downside of this approach, however, is that constructing valid confidence intervals becomes more challenging.

@@

\subsubsection{Notation}
 All notation introduced here pertains to the Gaussian model with intercepts (\ref{theoretical_model}). Recall that the attenuation function $\gamma: \R^4 \to \R$ is defined by
$$ \gamma(\beta^g_1, c, \pi, \beta^g_0) = \frac{\pi(\omega - \E[\hat{p}_i])}{ \E[\hat{p}_i](1 -\E[\hat{p}_i])},$$ where $$\begin{cases}
\omega = 1 - \Phi\left(c - \beta_1^g - \beta_0^g \right) ,\\ \zeta = 1 - \Phi\left( c - \beta^g_0 \right), \\ \E[\hat{p}_i] = \zeta(1-\pi) + \omega\pi. 
\end{cases}$$ Additionally, recall that the bias function $b: \R^5 \to \R$ is defined by
$$ b(\beta^g_1, c, \pi, \beta^g_0, \beta^m_1) = \beta^m_1 (1 - \gamma(\beta^g_1, c, \pi, \beta^g_0)).$$ The parameters $\beta^g_0$ and $\beta^m_1$ will be given, fixed constants throughout the proofs. Therefore, to minimize notation, we typically use $\gamma(\beta^g_1, c, \pi)$ (resp., $b(\beta^g_1, c, \pi)$) to refer to the function $\gamma$ (resp., $b$) evaluated at $(\beta^g_1, c, \pi, \beta^g_0)$ (resp., $(\beta^g_1, c, \pi, \beta^g_0, \beta^m_1)$). Next, we define the functions $g$ and $h: \R^4 \to \R$ by
\begin{equation}\label{def_g}
g(\beta^g_1, c, \pi, \beta^g_0) = (1-\pi)\left( \int_{ -\infty}^{\beta_0^g + \beta_1^g - c} f \right) - (1-\pi)\left( \int_{-\infty}^{ \beta_0^g - c} f \right)\end{equation}
 and
\begin{multline}\label{def_h}
h(\beta^g_1, c, \pi, \beta^g_0) = \left[(1-\pi)\left( \int_{-\infty}^{ \beta_0^g - c} f \right) + \pi\left( \int_{-\infty}^{ \beta^g_0 + \beta^g_1 - c} f \right) \right] \cdot \\ \left[(1-\pi)\left( \int_{-\infty}^{c - \beta^g_0} f \right) + \pi\left(\int_{-\infty}^{c - \beta_0^g - \beta_1^g} f \right) \right],
\end{multline} where $f: \R \to [0,\infty)$ is the $N(0,1)$ density. We will refer to $g$ and $h$ repeatedly throughout the proofs. Similarly to above, we typically use the notation $g(\beta^g_1, c, \pi)$ (resp.\ $h(\beta^g_1, c, \pi)$) to refer to $g$ (resp.\ $h$) evaluated at $(\beta^g_1, c, \pi, \beta^g_0)$.

@@

\subsubsection{Re-expressing $\gamma$ in a simpler form}

>>>>

\begin{document}
\noindent
Tim, Gene, Kathryn
\begin{center}
\textbf{CRISPR genome editing, single-cell sequencing, and exponential family measurement error models}
\end{center}

\begin{abstract}
CRISPR genome editing single-cell sequencing have transformed biological discovery. Single-cell CRISPR screens unite these two technologies, linking genetic perturbations in individual cells to changes in gene expression and illuminating regulatory networks underlying diseases. In this work we study single-cell CRISPR screens from a statistical perspective. First, we demonstrate on real data that a standard method for estimation and inference in single-cell CRISPR screens — “thresholded regression” — exhibits attenuation bias and a bias-variance tradeoff as a function of an intrinsic tuning parameter. We recover these phenomena in precise theoretical terms in an idealized Gaussian setting. Next, we introduce GLM-EIV (“generalized linear model with errors-in-variables”), a new method for single-cell CRISPR screen analysis. GLM-EIV generalizes the classical errors-in-variables model to response distributions and sources of measurement error that are exponential family-distributed, overcoming limitations of thresholded regression. We develop a computational infrastructure to deploy GLM-EIV across hundreds or thousands of processors on clouds (e.g., Microsoft Azure) and high-performance clusters. Leveraging this infrastructure, we apply GLM-EIV to analyze two recent, large-scale, single-cell CRISPR screen datasets, yielding new biological insights.
\end{abstract}

@@

\begin{equation}\label{theoretical_model}
\begin{cases}
m_i = \beta^m_0 + \beta^m_1 p_i + \ep_i \\
g_i = \beta^g_0 + \beta^g_1 p_i + \tau_i \\
p_i \sim \textrm{Bern}(\pi) \\
\ep_i, \tau_i \sim N(0,1) \\
p_i \indep \tau_i \indep \ep_i.
\end{cases}
\end{equation}
For a given threshold $c \in \R$, the imputed perturbation assignment $\hat{p}_i$ is $\hat{p}_i = \mathbb{I}(g_i \geq c).$ The thresholding estimator $\hat{\beta}^m_1$ for $\beta^m_1$ is given by $$\hat{\beta}^m_1 = \frac{\sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}}) (m_i - \overline{m})}{\sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}})}.$$

\begin{proposition}
 The almost sure limit (as $n \to \infty$) of $\hat{\beta}^m_1$ is
\begin{equation}\label{thresh_est_intercepts}
\hat{\beta}^m_1 \xrightarrow{a.s.} \beta^m_1 \left(\frac{ \pi( \omega - \E[ \hat{p}_i ])}{ \E[\hat{p}_i] (1 - \E[\hat{p}_i])}\right),
\end{equation}
where
$$\begin{cases}
\E[\hat{p}_i] = \zeta(1-\pi) + \omega\pi , \\
\omega = \Phi\left(\beta_1^g + \beta_0^g -c \right) ,\\ \zeta = \Phi\left( \beta^g_0 - c \right). \\
\end{cases}$$
\end{proposition}
Let $\gamma: \R^4 \to \R$ be defined by
$$ \gamma(\beta^g_1, \pi, c, \beta^g_0) = \frac{\pi (\omega - \E[\hat{p}_i])}{\E[\hat{p}_i] (1 - \E[\hat{p}_i]) }.$$ We call $\gamma$ the ``attenuation function.'' Observe that \begin{itemize}
\item[i.] $\gamma$ does not depend on $\beta^m_1$ or $\beta^m_0$, and
\item[ii.] $\hat{\beta}^m_1 \xrightarrow{a.s.} [\gamma(\beta_0^g, \beta_1^g, c, \pi)] \beta^m_1.$
\end{itemize}
Let $b: \R^5 \to \R$ be the asymptotic bias of $\hat{\beta}^m_1$, i.e.
$$
b(\beta^g_1, \pi, c, \beta^g_0, \beta^m_1) = \lim_{n\to\infty} \left(\beta^m_1 - \E[\hat{\beta}^m_1]\right) = \beta^m_1 - \E \left( \lim_{a.s.} \hat{\beta}^m_1 \right) = \beta^m_1 - \lim_{a.s.} \hat{\beta}^m_1,
$$
where the $\lim_{a.s.}$ denotes a.s.\ convergence. \textcolor{red}{(HOW TO MAKE THE ABOVE RIGOROUS?)} We can express $b$ in terms of $\gamma$ as $$b(\beta^g_1, \pi, c, \beta^g_0, \beta^m_1) = \beta^m_1 - \gamma(\beta^g_1, \pi, c, \beta^g_0)  \beta^m_1 = \beta^m_1 \left[1 - \gamma(\beta^g_1, \pi, c, \beta^g_0)  \right].$$ The asymptotic bias vanishes when the attenuation function equals $1$.

@@

To investigate the basic question of ``What is a good threshold selection strategy?'', we study the relationship between the asymptotic bias $b$ of $\hat{\beta}^m_1$ and the selected threshold $c$. For simplicity, we set the perturbation probability $\pi$ to $1/2$. Let $c_\textrm{bayes} \in \R$ be the Bayes-optimal decision boundary for classifying cells as perturbed or unperturbed, i.e. $$c_\textrm{bayes} = \argmin_{c \in \R} \P(\hat{p}_i \neq p_i).$$ Simple algebra shows that $c_\textrm{bayes} = \beta_0^g + (1/2) \beta^g_1.$ Below, we state several results on the limiting bias $b$ of $\hat{\beta}^m_1$. We refer throughout to Figure \ref{thresholding_theoretical}a, which displays plots of asymptotic bias versus threshold for different values of $\beta^g_1$; the target of inference, $\beta^m_1$, is fixed to $1$ in these plots.

\begin{itemize}
\item 
\begin{proposition}\label{prop:att_bias} Assume without loss of generality that $\beta^m_1 > 0.$ Fix $\pi = 1/2$. For all $(\beta^g_1, c, \beta^g_0) \in \R^3$,
	$$b(\beta^g_1, 1/2, c, \beta^g_0, \beta^m_1) > 0.$$
\end{proposition}
The thresholding method suffers strict attenuation bias for all choices of the threshold and over all possible model parameters (Figure \ref{thresholding_theoretical}a). Attenuation bias is a common attribute of estimators that ignore measurement in errors-in-variables models \cite{Stefanski2000a}.

\item \begin{proposition}
For $\pi = 1/2$ and given $(\beta^g_1, \beta^g_0, \beta^m_1) \in \R^3$, the Bayes-optimal decision boundary $c_\textrm{bayes}$ is a critical value of the limiting bias function $b$, i.e.
$$ \frac{\partial b}{\partial c}\left(\beta^g_1, 1/2, c_\textrm{bayes}, \beta^g_0, \beta^m_1\right) = 0.$$
\end{proposition}
The Bayes-optimal decision boundary is an optimum (or possibly a saddle point) of the limiting bias function (Figure \ref{thresholding_theoretical}a, vertical blue lines). Interestingly, $c_\textrm{bayes}$ is in some cases a maximizer of the bias (Figure \ref{thresholding_theoretical}a, left) and in other cases a minimizer of the bias (Figure \ref{thresholding_theoretical}a, right).
\item
\begin{proposition}\label{prop:limit}
Fix $\pi = 1/2$. As the threshold $c$ tends to infinity, the asymptotic bias $b$ tends to $\beta_1^m/2$, i.e.
$$\lim_{c \to \infty} b(\beta^g_1, 1/2, c, \beta^g_0, \beta^m_1) = \beta_1^m/2.$$
\end{proposition}
We always can set the threshold to a large number and attain a bias of $\beta_1^m/2$ (Figure \ref{thresholding_theoretical}a, all panels). This result establishes an upper bound on the bias of thresholded regression (under optimal threshold selection strategy).

 \item The following proposition compares the two threshold selection strategies introduced above (i.e., large number versus Bayes-optimal decision boundary) head-to-head.
 \begin{proposition} Let $\beta^m_1 > 0$. For $\beta^g_1 \in [0, 2\Phi^{-1}(3/4))$, we have that $$b(\beta^g_1, 1/2, c_\textrm{bayes}, \beta^g_0, \beta^m_1) > b(\beta^g_1, 1/2, \infty, \beta^g_0, \beta^m_1).$$ For $\beta^g_1 = 2\Phi^{-1}(3/4)$, we have that $$ b(\beta^g_1, 1/2, c_\textrm{bayes}, \beta^g_0, \beta^m_1) = b(\beta^g_1, 1/2, \infty, \beta^g_0, \beta^m_1).$$ Finally, for $\beta^g_1 \in (2\Phi^{-1}(3/4), \infty)$, we have that
 $$b(\beta^g_1, 1/2, c_\textrm{bayes}, \beta^g_0, \beta^m_1) < b(\beta^g_1, 1/2, \infty, \beta^g_0, \beta^m_1).$$
 \end{proposition}
In other words, setting the threshold to a large number yields a smaller bias when $\beta^g_1$ is small (i.e., $\beta^g_1 < 2\Phi^{-1}(3/4) \approx 1.35$; Figure \ref{thresholding_theoretical}a, left); setting the threshold to the Bayes-optimal decision boundary yields a smaller bias when $\beta^g_1$ is big (i.e., $\beta^g_1 > 2\Phi^{-1}(3/4)$; Figure \ref{thresholding_theoretical}a, right); and the two approaches coincide when $\beta^g_1$ is intermediate (i.e., $\beta^g_1 = 2\Phi^{-1}(3/4)$; Figure \ref{thresholding_theoretical}a, middle).

\item
 \begin{proposition}\label{prop_monotonic} Fix $\pi = 1/2$. The limiting bias $b$ decreases monotonically in $\beta_1^g$, i.e.
$$\frac{\partial b}{\partial(\beta^g_1)}\left(\beta^g_1, 1/2, c, \beta^g_0, \beta^m_1\right) \leq 0.$$
\end{proposition}
This result formalizes the intuition that the problem becomes easier as the gRNA mixture distribution becomes increasingly well-separated. To visualize Proposition (\ref{prop_monotonic}), one can fix a threshold (e.g., $c = 0$) and scan for bias across the panels.
\end{itemize}

These results are subtle, but we can summarize them as follows. First, selecting a threshold that minimizes the bias is challenging, as there is no rule of thumb that we can apply universally (e.g., ``always choose the Bayes-optimal decision boundary'' or ``always choose a large number'') due to the complexity of the bias function. Second, even if we \textit{have} selected a good threshold, we incur nonzero attenuation bias.

\begin{center}
\textbf{Generalizing to $\pi \in [0,1/2]$ (Panel b)}
\end{center}
We generalize the expression for bias to arbitrary $\pi \in [0,1/2]$:
\begin{proposition}
As the threshold $c$ tends to infinity, the limiting bias $b$ tends to $\beta^m_1 \pi,$ i.e.
$$ \lim_{ c \to \infty } b(\beta^g_1, \pi, c, \beta^g_0, \beta^m_1) = \beta^m_1 \pi.$$
\end{proposition}
That is, if the perturbation probability is $\pi$, and if we set the threshold to a large number, then the asymptotic bias is $\beta^m_1 \pi$ (Figure \ref{thresholding_theoretical}b; $\beta^m_1$ fixed at $1$). We can understand this result intuitively by considering an extreme example: when $\pi$ is very small (e.g., $\pi = 0.01$), almost all cells are unperturbed. Therefore, in selecting a large threshold, we correctly classify nearly all unperturbed cells as unperturbed; on the other hand, the \textit{perturbed} cells that we misclassify as \textit{unperturbed} are swamped in number by the truly unperturbed cells, resulting in a small bias.

\begin{figure}[h!]
	\centering
	\includegraphics[width=1\linewidth]{../../figures/thresholding_theoretical/plot}
	\caption{\textbf{Theoretical challenges of thresholded regression.} \textbf{a,} (Asymptotic) bias versus threshold for different values of $\beta^g_1$. The bias function is highly nonconvex and strictly nonzero. Across all panels, $\beta^g_0 = 0, \beta^m_1 = 1$, and $\pi = 1/2$. \textbf{b,} Bias versus $\pi$ when the threshold is set to a large number. Bias is directly proportional to $\pi$ in this setting. \textbf{c,} Bias-variance decomposition of thresholding estimator in no-intercept model. Bias decreases and variance increases as the threshold tends to infinity. $\beta^m_1 = 1, \beta^g_1 = 1,$ and $\pi = 0.1$.}
	\label{thresholding_theoretical}
\end{figure}

@@

Finally, to shed light on limitations of the large threshold selection strategy, we derive an exact bias-variance decomposition for the thresholding estimator. We consider a slightly simpler, no-intercept version of (\ref{theoretical_model}) for this purpose:
\begin{equation}\label{theoretical_model_no_int}
\begin{cases}
m_i = \beta_m p_i + \ep_i \\
g_i = \beta_g p_i + \tau_i \\
p_i \sim \textrm{Bern}(\pi) \\
\ep_i, \tau_i \sim N(0,1) \\
p_i \indep \tau_i \indep \ep_i.
\end{cases}
\end{equation}
The thresholding estimator $\hat{\beta}_m$ in the no-intercept case is
\begin{equation}\label{thresh_estimator_no_int}
 \hat{\beta}_m = \frac{ \sum_{i=1}^n \hat{p}_i m_i }{\sum_{i=1}^n \hat{p}_i^2}. 
\end{equation}
\begin{proposition}
The limiting distribution of $\hat{\beta}_m$ is
$$\sqrt{n}(\hat{\beta}_m - l) \xrightarrow{d} N\left(0, \frac{ \beta_m\omega\pi(\beta_m - 2l) + \E[\hat{p}_i](1 + l^2) }{\left(\E[\hat{p}_i]\right)^2} \right),$$ where
$$\begin{cases}
l = \beta_m\omega \pi/[\zeta(1-\pi) + \omega \pi], \\
\E[\hat{p}_i] = \pi \omega + (1-\pi) \zeta, \\
\omega = \Phi(\beta_g - c), \\
\zeta = \Phi(- c).\\
\end{cases}
$$
\end{proposition}
This result yields an exact bias-variance decomposition for $\hat{\beta}_m$ for large $n$ (Figure \ref{thresholding_theoretical}c). As the threshold tends to infinity, the bias decreases and the variance increases, consistent with the intuition that a large threshold reduces the misclassification rate at the cost of decreasing the ``effective sample size.'' The best strategy for maximizing estimation accuracy (as quantified by mean squared error) is to select a threshold that induces moderate bias. A downside of this approach, however, is that constructing valid confidence intervals becomes more challenging.

@@

\subsubsection{Notation}
 All notation introduced here pertains to the Gaussian model with intercepts (\ref{theoretical_model}). Recall that the attenuation function $\gamma: \R^4 \to \R$ is defined by
$$ \gamma(\beta^g_1, c, \pi, \beta^g_0) = \frac{\pi(\omega - \E[\hat{p}_i])}{ \E[\hat{p}_i](1 -\E[\hat{p}_i])},$$ where $$\begin{cases}
\omega = \Phi\left(c - \beta_1^g - \beta_0^g \right) ,\\ \zeta = \Phi\left( c - \beta^g_0 \right), \\ \E[\hat{p}_i] = \zeta(1-\pi) + \omega\pi. 
\end{cases}$$ Additionally, recall that the bias function $b: \R^5 \to \R$ is defined by
$$ b(\beta^g_1, c, \pi, \beta^g_0, \beta^m_1) = \beta^m_1 (1 - \gamma(\beta^g_1, c, \pi, \beta^g_0)).$$ The parameters $\beta^g_0$ and $\beta^m_1$ are given, fixed constants throughout the proofs. Therefore, to minimize notation, we typically use $\gamma(\beta^g_1, c, \pi)$ (resp., $b(\beta^g_1, c, \pi)$) to refer to the function $\gamma$ (resp., $b$) evaluated at $(\beta^g_1, c, \pi, \beta^g_0)$ (resp., $(\beta^g_1, c, \pi, \beta^g_0, \beta^m_1)$). Next, we define the functions $g$ and $h: \R^4 \to \R$ by
\begin{equation}\label{def_g}
g(\beta^g_1, c, \pi, \beta^g_0) = (1-\pi)\left( \int_{ -\infty}^{\beta_0^g + \beta_1^g - c} f \right) - (1-\pi)\left( \int_{-\infty}^{ \beta_0^g - c} f \right)\end{equation}
 and
\begin{multline}\label{def_h}
h(\beta^g_1, c, \pi, \beta^g_0) = \left[(1-\pi)\left( \int_{-\infty}^{ \beta_0^g - c} f \right) + \pi\left( \int_{-\infty}^{ \beta^g_0 + \beta^g_1 - c} f \right) \right] \cdot \\ \left[(1-\pi)\left( \int_{-\infty}^{c - \beta^g_0} f \right) + \pi\left(\int_{-\infty}^{c - \beta_0^g - \beta_1^g} f \right) \right],
\end{multline} where $f: \R \to [0,\infty)$ is the $N(0,1)$ density. We will refer to $g$ and $h$ repeatedly throughout the proofs. Similarly to above, we typically use the notation $g(\beta^g_1, c, \pi)$ (resp.\ $h(\beta^g_1, c, \pi)$) to refer to $g$ (resp.\ $h$) evaluated at $(\beta^g_1, c, \pi, \beta^g_0)$.

@@

\subsubsection{Re-expressing $\gamma$ in a simpler form}
We rewrite the attenuation fraction $\gamma$ in a way that makes it more amenable to theoretical analysis. We leverage the fact that $f$ integrates to unity and is even.
First,
\begin{multline}\label{thm:gamma_expression_1} \E\left[\hat{p}_i\right] = (1 - \pi) \left( \int_{c - \beta^g_0}^\infty f \right) + \pi \left(\int_{c - \beta_0^g - \beta_1^g}^\infty f \right) \\ = (1 - \pi) \left( \int_{-\infty}^{ \beta_0^g - c} f \right) + \pi \left( \int_{-\infty}^{ \beta^g_0 + \beta^g_1 - c} f \right), \end{multline}
 and so \begin{multline}\label{thm:gamma_expression_2} 1 - \E\left[\hat{p}_i\right] = (1 - \pi) + \pi - \E[\hat{p}_i]  = (1-\pi) \left(1 - \int_{c - \beta_0^g}^\infty f \right)  + \pi \left(1 - \int_{c - \beta^g_0 - \beta^g_1}^\infty f \right) \\ = (1 - \pi) \left( \int_{-\infty}^{c - \beta^g_0} f \right) + \pi \left( \int_{-\infty}^{c - \beta_0^g - \beta_1^g} f \right).
\end{multline}
Next,
\begin{equation}\label{thm:gamma_expression_3}
\omega = \int_{c - \beta^g_1 - \beta^g_0}^\infty f,\end{equation} and so
\begin{multline}\label{thm:gamma_expression_4}
\omega - \E[\hat{p}_i]  = \left( \int_{c - \beta^g_1 - \beta^g_0}^\infty f \right) - (1 - \pi) \left( \int_{c - \beta^g_0}^\infty f \right) - \pi \left( \int_{c - \beta_0^g - \beta_1^g}^\infty f \right) \\ = (1 - \pi) \left(\int_{c - \beta^g_0 - \beta^g_1 }^\infty f \right) - (1- \pi) \left(\int_{c - \beta^g_0}^\infty f \right) \\ =  (1 - \pi)\left(\int_{-\infty}^{\beta_0^g + \beta_1^g - c} f \right) - (1 - \pi) \left( \int_{-\infty}^{ \beta_0^g - c} f \right).
\end{multline}
Combining (\ref{thm:gamma_expression_1}, \ref{thm:gamma_expression_2}, \ref{thm:gamma_expression_3}, \ref{thm:gamma_expression_4}), we find that
\begin{multline}\label{gamma_alternative}
\gamma(\beta^g_1, c, \pi) = \frac{\pi(\omega - \E[\hat{p}_i])}{\E[\hat{p}_i](1 - \E[\hat{p}_i])} \\ = \frac{\pi \left[ (1 - \pi) \left( \int_{ -\infty}^{\beta_0^g + \beta_1^g - c} f \right) - (1 - \pi) \left( \int_{-\infty}^{ \beta_0^g - c} f \right) \right]}{\left[(1-\pi) \left( \int_{-\infty}^{ \beta_0^g - c} f \right) + \pi \left( \int_{-\infty}^{ \beta^g_0 + \beta^g_1 - c} f \right) \right] \left[(1 - \pi) \left( \int_{-\infty}^{c - \beta^g_0} f \right) + \pi \left( \int_{-\infty}^{c - \beta_0^g - \beta_1^g} f \right) \right]}.
\end{multline}
As a corollary, when $\pi = 1/2$,
\begin{multline}\label{gamma_alternative_pi_half}
\gamma(\beta^g_1, c, 1/2)  \\ = \frac{\left( \int_{ -\infty}^{\beta_0^g + \beta_1^g - c} f \right) - \left( \int_{-\infty}^{ \beta_0^g - c} f \right)}{\left[ \left( \int_{-\infty}^{ \beta_0^g - c} f \right) + \left( \int_{-\infty}^{ \beta^g_0 + \beta^g_1 - c} f \right) \right] \left[\left( \int_{-\infty}^{c - \beta^g_0} f \right) + \left(\int_{-\infty}^{c - \beta_0^g - \beta_1^g} f \right) \right]}.
\end{multline}
Recalling the definitions of $g$ (\ref{def_g}) and $h$ (\ref{def_h}), we can write $\gamma$ as
$$ \gamma(\beta^g_1, c, \pi) = \frac{\pi g(\beta^g_1, c, \pi)}{h(\beta^g_1, c,\pi)}.$$
The special case (\ref{gamma_alternative_pi_half}) is identical to
\begin{equation}\label{gamma_alt2_pi_half}
\gamma(\beta^g_1, c, 1/2) = \frac{(4)(1/2)g(\beta^g_1, c, 1/2)}{4 h(\beta^g_1, c, 1/2)} = \frac{2 g(\beta^g_1, c, 1/2)}{4h(\beta^g_1, c, 1/2)},
\end{equation}
i.e., the numerator and denominator of  (\ref{gamma_alt2_pi_half}) coincide with those of (\ref{gamma_alternative_pi_half}).
